---
title: "Distribution on the amount of trash collected by Mr. Trash Wheel in 2023"
subtitle: "Proposal"
author: 
  - name: "Iris Sum"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "This project review the distribution of trash collected by Mr. Trash Wheel in 2023."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
    code-fold: true
editor: visual
code-annotations: hover
execute:
  warning: false
  message: false
  error: false
---

```{r}
#| label: load-pkgs

if (!require("pacman")) 
  install.packages("pacman")
  
pacman::p_load(here,
               tidytuesdayR,
               tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| results: false

rm(list=ls())
dffile <- tidytuesdayR::tt_load('2024-03-05')

df <- dffile$trashwheel # load data set
```

The `trashwheel` dataset provided by Mr. Trash Wheel, one of the four Trash Wheels in the Healthy Harbor initiative, was selected from Tidy Tuesday 2024 for this project. It records the data of the trash collected by the four semi-autonomous trash interceptors placed at the end of a river, stream or other outfall, namely the *Mister Trash Wheel*, *the Professor Trash Wheel*, *the Captain Trash Wheel* and *the Gwynnda Trash Wheel* from 2014 to 2023.

Since the trash wheel started operating from 2014, there were nearly 2,362 tons of trash collected, which equal to the weight of 15 whales! Among those collected trash, there were over 13 millions of cigarette butts. It really helps saving our ocean from toxic plastic pollution.

There are 16 variables with 993 records in the dataset. Each observation represents a dumpster from one of the four trash interceptors, and its related values such as the date of collection, weight, volume and the number of different trash collected. In total 7 types of trash are recorded as attributes and the collection amount are stored as numerical data: `PlasticBottles`, `Polystyrene`, `CigaretteButts`, `GlassBottles`, `PlasticBags`, `Wrappers`, `SportsBalls`.

The selection of dataset was based on its cleanliness, simplicity, and meaningfulness. It is astonishing when looking into the amount of trash that has been disposed into rivers and the impressive collection capacity of the trash wheels. The objective of this project is to present the dataset in a clean and eye-catching manner in order to engage individuals in the issue of trash pollution and appreciate the endeavors dedicated to preserving the environment from trash.

## Questions

What is the distribution of trash collected by Mr. Trash Wheel in 2023?

## Analysis plan

This project will review the distribution of trash collected by Mr. Trash Wheel in 2023 from two aspects: by the types of the trash collected and by the trash interceptors. The data wrangling will mainly rely on `tidyverse` library.

Attributes used for the project (filtered by 2023):

```{r}
#| label: attribute-selection

df <- df |>
  filter(Year == 2023) |> # filter the dataset in 2023
  select(Name, Month, PlasticBottles, Polystyrene, CigaretteButts, GlassBottles, PlasticBags, Wrappers, SportsBalls) # select related atributes for analysis

glimpse(df) # take a look of the filtered dataset
```

#### Plot 1:

From types of the trash collected perspective, separate plots by `geom_pictogram()` will be created for each trash interceptor. It aims to visualized the proportion of differ trash collected by each trash interceptor clearly for comparison, with the use of icons to represent different trash types instead of single image with `geom_waffle()`.

The variables involved in this plot will be `Name` containing the name of interceptors, and the 7 trash type variables. The dataset will first be filtered by `Year` 2023, followed by transferring all trash column names into a new variable `trash_type` with `trash_amount` storing the numeric amount of each trash by using `pivot_longer()`. Then the data will be `group_by()` `Name` and `trash_type`, and calculate the sum of trash collected with `summarise()`.

#### Plot 2:

On the other hand, a plot by `geom_col()` with x-axis equals to `Month` and y-axis equal to `total_trash` will be created to show the change of trash amount collected by each trash interceptor with dodged columns throughout the year 2023. To do this, after filtering the dataset with `Year` 2023, variables `Name`, `Month`, and the 7 trash type variables will be selected. A new variable `total_trash` will be added by summing up all trash amount collected by each trash interceptor rowwisely.

For statistical analysis, ANOVA will be used for observing the variance between total trash amount by each trash interceptor throughout the year. To prepare the data matrix for this, the cleaned data from plot 2 will be used by removing `Month` and the 7 trash type variables, remaining `Name` and `total_trash`. ANOVA will be performed with `aov()` and the result will be interpreted in Discussion section with information from plot 2.

### Plan of Attack:

+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
| Task Name                          | Status     | Due     | Priority | Summary                                                               |
+:==================================:+:==========:+:=======:+:========:+:=====================================================================:+
| Draft of proposal                  | Complete   | 26-May  | High     | Choose of dataset.                                                    |
|                                    |            |         |          |                                                                       |
|                                    |            |         |          | Decide project questions and variables used.                          |
+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
| Finalize project outline           | Complete   | 3-Jun   | High     | Amend the `proposal.qmd` with feedbacks from classmates and professor |
+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
| Complete waffle plot for part 1    | Incomplete | 15-Jun  | Moderate | Complete waffle plot and its interpretation                           |
+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
| Complete bar chart for part 2      | Incomplete | 18-Jun  | Moderate | Complete bar chart and its interpretation                             |
+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
| Finalize the report                | Incomplete | 20-Jun  | Low      | Finalize the `index.qmd`                                              |
+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
| Finalize the presentable materials | Incomplete | 20-Jun  | Low      | Finalize the `presentation.qmd`                                       |
+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
| Record the presentation            | Incomplete | 24-Jun  | Low      | Complete the presentation                                             |
+------------------------------------+------------+---------+----------+-----------------------------------------------------------------------+
